Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2020-01-10 draft.tex
New draft section

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{newunicodechar}
\usepackage{textcomp}
%\newunicodechar{⇒}{\implies}
\newunicodechar{⇒}{\Rightarrow}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≤}{\leq}
\newunicodechar{≥}{\geq}
%… Horizontal Ellipsis
\DeclareUnicodeCharacter{2026}{\ifmmode\dots\else\textellipsis\fi}
\newunicodechar{∧}{\land}
\newunicodechar{∨}{\lor}
\newunicodechar{∩}{\cap}
\newunicodechar{∪}{\cup}
%¬ Not Sign
\DeclareUnicodeCharacter{00AC}{\ifmmode\lnot\else\textlnot\fi}
\newunicodechar{⇔}{\Leftrightarrow}
\newcommand{\N}{ℕ}
\newunicodechar{ℕ}{\mathbb{N}}
\usepackage{algorithm, algpseudocode}
\usepackage{amsmath,amssymb,enumerate,amsthm}
\usepackage{natbib}
\usepackage[strict]{siunitx}
\usepackage{hyperref}
\usepackage{mathrsfs}

@@

\DeclareMathOperator{\Regret}{Regret}
\DeclareMathOperator{\SCORE}{Score}
\DeclareMathOperator{\PMR}{PMR}
\DeclareMathOperator{\MR}{MR}
\DeclareMathOperator{\MMR}{MMR}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

@@

Observe that if $\MMR(\pprofile, W)\!=\!0$, then $x^{*}_{\pprofile,W}$ is a necessary co-winner; this means that for any valid completion of the profile and any feasible $w \!\in\! W$, $x^{*}_{\pprofile,W}$ obtains a highest score.

@@

\appendix
\section{Bounding differences of weights}
We want to check whether we can round PMR values to (for example) $10^{-4}$. Thus, we want to bound the difference of scores (unless it is zero).

@@

Problem: we can make the difference of scores very small, even with $n=2$ voters and $m=3$ alternatives and while satisfying convexity. Consider $w_1 = 1, w_2 = \frac{1-\epsilon}{n}, w_3 = 0$, $s(a) - s(b) = (w_1 - w_2) - (n-1) (w_2 - w_3) = w_1 - n w_2 = \epsilon$. Choose $\epsilon < \frac{1}{n+1}$ so that $w_2 > \epsilon$ and $n ≥ 2$ to satisfy convexity. For example, $\epsilon = 1/1000, n = 2, w_2 = 999/2000$.
\end{document}

>>>>

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{newunicodechar}
\usepackage{textcomp}

%\newunicodechar{⇒}{\implies}
\newunicodechar{⇒}{\Rightarrow}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≤}{\leq}
\newunicodechar{≥}{\geq}
%… Horizontal Ellipsis
\DeclareUnicodeCharacter{2026}{\ifmmode\dots\else\textellipsis\fi}
\newunicodechar{∧}{\land}
\newunicodechar{∨}{\lor}
\newunicodechar{∩}{\cap}
\newunicodechar{∪}{\cup}
%¬ Not Sign
\DeclareUnicodeCharacter{00AC}{\ifmmode\lnot\else\textlnot\fi}
\newunicodechar{⇔}{\Leftrightarrow}
\newcommand{\N}{ℕ}
\newunicodechar{ℕ}{\mathbb{N}}
\newcommand{\R}{ℝ}
\newunicodechar{ℝ}{\mathbb{R}}
\usepackage{algorithm, algpseudocode}
\usepackage{amsmath,amssymb,enumerate,amsthm}
\usepackage{natbib}
\usepackage[strict]{siunitx}
\usepackage{hyperref}
\usepackage{mathrsfs}

@@

\DeclareMathOperator{\Regret}{Regret}
\DeclareMathOperator{\SCORE}{Score}
\DeclareMathOperator{\PMR}{PMR}
\DeclareMathOperator{\MR}{MR}
\DeclareMathOperator{\MMR}{MMR}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\badn}{badness}

@@

Observe that if $\MMR(\pprofile, W)\!=\!0$, then $x^{*}_{\pprofile,W}$ is a necessary co-winner; this means that for any valid completion of the profile and any feasible $w \!\in\! W$, $x^{*}_{\pprofile,W}$ obtains the highest score.

@@

\appendix
\section{Bounding differences of weights}
\newcommand{\Wq}{W_\text{q}}

We want to check whether we can round PMR values to (for example) $10^{-4}$. Thus, we want to bound the difference of scores (unless it is zero).

@@

Problem: we can make the difference of scores very small, even with $n=2$ voters and $m=3$ alternatives and while satisfying convexity. Consider $w_1 = 1, w_2 = \frac{1-\epsilon}{n}, w_3 = 0$, $s(a) - s(b) = (w_1 - w_2) - (n-1) (w_2 - w_3) = w_1 - n w_2 = \epsilon$. Choose $\epsilon < \frac{1}{n+1}$ so that $w_2 > \epsilon$ and $n ≥ 2$ to satisfy convexity. For example, $\epsilon = 1/1000, n = 2, w_2 = 999/2000$.

\section{Best question}
\subsection{Much fixed}
Throughout, $\Wq$ designates a closed set.

Fix $x, y, \profile^*, W \subseteq \W$, fix $m = 3$. $\Wq \subseteq W$, define $\PMR(\Wq) = \PMR(x, y, \profile^*, \Wq)$. Define $\PMR(w) = PMR(\set{w})$, denoting the pairwise regret when the weight vector is $w$. Let $P$ be a probability defined on the whole set $\W$ of possible weights.
%Define $\overline{w} = \argmax_{w \in W} s(y, \profile^*, w) - s(x, \profile^*, w) = \argmax PMR(w)$.

For $\emptyset ≠ \Wq \subset W$, define $\badn(\Wq) = P[w^* \in \Wq] \PMR(\Wq) + P[w^* \notin \Wq] \PMR(W \setminus \Wq).$
Equivalently, we could define the quality of a question as $P[w^* \in \Wq] (PMR(W) - PMR(\Wq)) + P[w^* \notin \Wq] (PMR(W) - PMR(W \setminus \Wq))$. This yields the same ordering. Still equivalently, we could define the badness using $P[\Wq \mid W]$ and $P[W \setminus \Wq \mid W]$, so that these probabilities sum to one.

NB in general, this badness measure could perhaps be too short sighted, but that doesn’t seem to be a problem with the restrictions adopted here.

With $m=3$, we can consider that $W \subseteq [0, 1]$, with $w \in W$ denoting the weight associated with the second rank.
Let $P$ be the Lebesgue measure (or the uniform probability distribution) on $[0, 1]$.

Recalling that $\alpha^y_1$ denotes the number of times $y$ is ranked first in $\profile^*$, define $t_1 = \alpha^y_1 - \alpha^x_1$ and $t_2 = \alpha^y_2 - \alpha^x_2$ to represent the difference in number of voters placing at a given position. Then $\PMR(w) = t_1 + t_2 w$. Thus, the worst (maximal) regret $\PMR(\Wq)$ is reached, for any set of weights $\Wq$, at an extremum of $\Wq$. If $t_2 ≥ 0$, the supremum of $\Wq$ is an optimal adversarial choice, and if $t_2 ≤ 0$, the infimum of $\Wq$ is an optimal adversarial choice. Thus $\PMR(\Wq) = t_1 + t_2 \sup(\Wq)$ if $t_2 ≥ 0$ and $\PMR(\Wq) = t_1 + t_2 \inf(\Wq)$ if $t_2 ≤ 0$.

Assume that $W$ has the form $W = [\inf W, \sup W]$.

Let $W_q$, with $q \in W$, denote asking whether $w^* ≥ q$, thus $W_q = \set{w ≥ q} ∩ W$. 
We can say the following about the question $W_q$.
\begin{itemize}
	\item $P[W_q] = \sup W - q$
	\item $P[W \setminus W_q] = q - \inf W$
	\item $\PMR(W_q) = t_1 + t_2 \sup W$ if $t_2 ≥ 0$
	\item $\PMR(W_q) = t_1 + t_2 q$ if $t_2 ≤ 0$
	\item $\PMR(W \setminus W_q) = t_1 + t_2 q$ if $t_2 ≥ 0$. 
	\item $\PMR(W \setminus W_q) = t_1 + t_2 \inf W$ if $t_2 ≤ 0$
	\item $\badn(W_q) = (\sup W - q) (t_1 + t_2 \sup W) + (q - \inf W) (t_1 + t_2 q)$ if $t_2 ≥ 0$, whose derivative is $-(t_1 + t_2 \sup W) + (t_1 + t_2 q) + (q - \inf W) t_2 = 2 t_2 q - t_2 \sup W - t_2 \inf W$, thus the least badness is reached at $q = \frac{\sup W + \inf W}{2}$
	\item $\badn(W_q) = (\sup W - q) (t_1 + t_2 q) + (q - \inf W) (t_1 + t_2 \inf W)$ if $t_2 ≤ 0$, whose derivative is $-(t_1 + t_2 q) + (\sup W - q) t_2 + (t_1 + t_2 \inf W) = -2 t_2 q + t_2 \sup W + t_2 \inf W$, thus the least badness is reached at $q = \frac{\sup W + \inf W}{2}$
\end{itemize}
In both cases, asking a question about the middle of the interval is optimal.

\subsection{y variable}
Fix $x, \profile^*, W \subseteq \W$, fix $m = 3$. 
Assume that $W$ has the form $W = [\inf W, \sup W]$.
For $\Wq \subseteq W$, define $\MR(\Wq) = \max_{y \in A} \PMR(x, y, \profile^*, \Wq) = \max_{y \in A, w \in \Wq} s(y, \profile^*, w) - s(x, \profile^*, w)$. For $w \in W$, define $\MR(w) = \MR(\set{w})$.

For $\Wq \subseteq W$, define $\badn(\Wq) = P[\Wq] \MR(\Wq) + P[W \setminus \Wq] \MR(W \setminus \Wq)$.

Let $t_1^a = \alpha_1^a - \alpha_1^x$ and $t_2^a = \alpha_2^a - \alpha_2^x$, for $a \in A$, representing the difference of number of times an alternative reaches a given position compared to the alternative $x$. Thus, $t_1^x = t_2^x = 0$. For $q \in W$, $\PMR(x, a, \profile^*, q) = t_1^a + t_2^a q$. Thus, $\MR(q) = \max \set{t_1^x + t_2^x q, t_1^y + t_2^y q, t_1^z + t_2^z q}$, with $A = \set{x, y, z}$. Therefore, $\MR$ has a maximal value at an extremum point of $W$.

We obtain that $\MR(\Wq) = \max \set{\MR(\inf \Wq), \MR(\sup \Wq)}$.

Define $W_q = \set{w ≥ q} ∩ W$, with $q \in W$.

Assume that $t_2^y ≤ 0$ and $t_2^z ≤ 0$. 
Sort alternatives by slope, thus, let $a, b, c$ denote different alternatives such that $t_2^a ≤ t_2^b ≤ t_2^c$ and $t_2^a = t_2^b ⇒ t_1^a ≥ t_1^b$ and $t_2^b = t_2^c ⇒ t_1^b ≥ t_2^b$ (picking one naming arbitrarily if several namings are compatible with these constraints).
TODO it should follow that $\inf W$ is a maximum of $\MR$ over $W$. 
Let $q_{a \sim b} \in ℝ$ denote the point such that $t_1^a + t_2^a q_{a \sim b} = t_1^b + t_2^b q_{a \sim b}$. If several or no points match this requirement, define $q_{a \sim b} = -1$. Define similarly $q_{b \sim c}$; setting it at 1 if several or no points match the requirement.
TODO it follows that $q_{a \sim b} ≤ q_{b \sim c}$.

TODO it follows that $\MR(q) = t_1^a + t_2^a q$ for $q ≤ q_{a \sim b}$, 
$\MR(q) = t_1^b + t_2^b q$ for $q_{a \sim b} ≤ q ≤ q_{b \sim c}$, 
$\MR(q) = t_1^c + t_2^c q$ for $q_{b \sim c} ≤ q$.

Also, $\MR(W_q) = \MR(q)$, and 
$\MR(W \setminus W_q) = \MR(\inf W)$.

Thus for $q ≤ q_{a \sim b}$, $\badn(W_q) = (\sup W - q) (t_1^a + t_2^a q) + (q - \inf W) (t_1^a + t_2^a \inf W)$.

\begin{example}
	When $q_{a \sim b} ≤ \inf W$ and $q_{b \sim c} ≤ \frac{1}{2}$, $q_{b \sim c}$ is an optimum. (TODO prove it)
	\begin{itemize}
		\item Def $\inf W = 0, \sup W = 1$.
		\item Def $t_1^a = 0, t_2^a = -20, t_1^b = 1, t_2^b = -10$.
		\item $q_{a \sim b} = -1/10, q_{b \sim c} = 1/10$.
		\item $\MR(\frac{1}{20}) = \frac{1}{2}$, $\MR(\frac{1}{10}) = 0$.
		\item $\MR(\frac{3}{20}) = 0 = \MR(\frac{4}{20})$.
		\item $\MR(W \setminus W_q) = 1, \forall q \in W$.
		\item $\badn(W_\frac{1}{20}) = \frac{19}{20} \frac{1}{2} + \frac{1}{20} 1 = \frac{21}{40}$.
		\item $\badn(W_\frac{1}{10}) = \frac{9}{10} 0 + \frac{1}{10} 1 = \frac{4}{40}$.
		\item $\badn(W_q) = (1 - q) 0 + q 1 = q$, with $\frac{1}{10} ≤ q$.
	\end{itemize}
	The optimum is at $\frac{1}{10}$ as expected.
\end{example}

\begin{example}
	Here, the optimum should be for some $q_{a \sim b} ≤ q ≤ \frac{1}{2}$.
	\begin{itemize}
		\item Def $\inf W = 0, \sup W = 1$.
		\item Def $t_1^a = 1, t_2^a = -10, t_1^b = 8/15, t_2^b = -2/3$.
		\item $q_{a \sim b} = 1/20, q_{b \sim c} = 8/10$.
		\item $\MR(\frac{1}{40}) = \frac{3}{4}$, $\MR(\frac{1}{20}) = \frac{1}{2}$.
		\item $\MR(\frac{2}{20}) = \frac{14}{30}$, $\MR(\frac{3}{20}) = \frac{13}{30}$, $\MR(\frac{4}{20}) = \frac{12}{30}$, $\MR(\frac{6}{20}) = \frac{10}{30}$, $\MR(\frac{1}{2}) = \frac{3}{15}$.
		\item $\MR(W \setminus W_q) = 1, \forall q \in W$.
		\item $\badn(W_\frac{1}{40}) = \frac{39}{40} \frac{3}{4} + \frac{1}{40} 1 = \frac{3}{4} + \frac{1}{160}$.
		\item $\badn(W_\frac{1}{20}) = \frac{19}{20} \frac{1}{2} + \frac{1}{20} 1 = \frac{1}{2} + \frac{1}{40}$.
		\item $\badn(W_\frac{2}{20}) = \frac{9}{10} \frac{7}{15} + \frac{1}{10} 1 = \frac{1}{2} + \frac{1}{50}$.
		\item $\badn(W_\frac{3}{20}) = \frac{17}{20} \frac{13}{30} + \frac{3}{20} 1 = \frac{1}{2} + \frac{11}{600}$.
		\item $\badn(W_\frac{4}{20}) = \frac{8}{10} \frac{12}{30} + \frac{2}{10} 1 = \frac{1}{2} + \frac{1}{50}$.
		\item For $\frac{1}{20} ≤ q ≤ \frac{8}{10}$, $\badn(W_q) = (1 - q) (\frac{8}{15} - \frac{2}{3} q) + q$; derivative is $-\frac{1}{5} + \frac{4}{3} q$, zero when $q = \frac{3}{20}$.
	\end{itemize}
\end{example}

\subsection{x variable}
Define $\MR_x(\Wq) = \max_{y \in A, w \in W_q} \PMR(x, y, \profile^*, w)$. Define $\MMR(\Wq) = \min_{x \in A} \MR_x(\Wq)$. Define $\badn(W_q) = (1 - q) \MMR(W_q) + q \MMR(W \setminus \W_q$.

Define $\MR_x(w) = \MR_x(\set{w})$.

We want to 1. Plot $\MR_x(w) = \max \set{t_1^{a, x} + t_2^{a, x} w, …}$. 2. Plot $\MR_x(W_q)$. 3. Plot $\MMR(W_q) = \min_x \MR_x(W_q)$.
\end{document}
